\textbf{Teaching a machine to generate an artificial universe}

### 2020-08-13 report.tex
redo: abstract; add: test figures, sections

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents separate two- and three-dimensional generative frameworks that use Generative Adversarial 
		    Networks to create realistic samples of cosmological structure. Provisional samples of the virtual cosmic web 
		    are obtained from N-body simulation data that is restricted to volumes and projections of size $80 \, h^{-1} 
		    \, \text{Mpc}$. The generated samples are shown to demonstrate a high degree of coherence with their simulated
		    counterparts and the distinction cannot be made visually by human experts. The nature of the separation between
		    generated and N-body simulation samples is quantified with repeated tests of the Kullback-Leibler divergence on
		    ensembles of samples alongside samples with uniformly distributed pixel densities. It is found that the 
		    implementations secure a coherence to within the same order of magnitude as a self comparison of the simulation
		    samples. The main advantage of the generative approach to simulating large volumes of cosmological structure is 
		    that much less time is required than creating N-body simulations on high performance computing resources. 
		    Future explorations of the method that are relevant to the next paradigm in cosmological simulations are also 
		    outlined. The most pertinent application being for a generator that can produce samples that maintain high 
		    statistical coherence to simulation data at various redshifts. At this stage, the generative artificial 
		    intelligence can then be said to be capable of generating an artificial universe.

@@

\setcounter{tocdepth}{2}
\begingroup
\let\clearpage\relax
\tableofcontents
\endgroup
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Beyond the distance scales of galactic diameters, toward the sizes of galaxy clusters and beyond, matter in the Universe 
appears as a webbed structure of filaments across volumes of emptiness known as voids. This network of luminous formation 
is known as the cosmic web~\cite{bond_cw1, Coles_cw2, Forero_cw3, Dietrich_cw4, cw_phenomenon}. The cosmic web as a 
distribution of matter contains artefacts linked to the state of the primordial universe, the nature of early universe 
perturbations and more fundamental physics to describe this era in cosmic time. The emergence of the structure is from 
an early period of exponential expansion, known as inflation~\cite{inflation_cosmo2}. This structure has been observed by 
missions such as~\cite{sdss1, 2df} and by weak gravitational lensing~\cite{glens_book} through distortions, or 
shears~\cite{shear_review}, to images of distant galaxies caused by the large-scale structure of the universe.

@@

\subsection{Generative adversarial learning}
% introduce adversarial relation / game between D and G
The discriminator and the generator are neural networks and each network tries to reduce their own objective function. 
This adversarial relationship is a game between the models rather than two optimisation problems. The optimum in this 
case is a Nash equilibrium~\cite{NIPS16} rather than a local minimum given by optimisation in the case of a singular 
neural network. In this equilibrium the returns of the objective functions only change by indirect influence of the 
other network and at this stage the gradients for backpropagation to either model have vanished.

@@

\begin{equation}
    \bm{\theta}^{(G)*} = \argmin_{\mathbf{\theta}^{(G)}}\max_{\mathbf{\theta}^{(D)}}V({\mathbf{\theta}^{(D)}},{\mathbf{\theta}^{(G)}})
\end{equation}

@@

The cross-entropy for the discriminator classification of a sample against the true label of the sample is numerically 
effective for a GAN in practice because this objective does not saturate for an incorrect identification. Because the 
networks try to minimize their opposing objectives, if the discriminators performance increases enough, the generator 
objective function tends to zero. This causes the generator performance to become static. At all points in the parameter 
space the generator objective minimization is guided by the gradients from backpropagation which will reduce to zero. At 
this point the discriminator cannot improve and less information is back-propagated through the networks.

@@

@@

\subsection{Gravitational wave sources} % Rachel Gray paper
Measurements of cosmological parameters from observations of gravitational wave events are independent of the 
electromagnetic (EM) spectrum and any obstacles to the transmission of EM radiation. Because these measurements 
do not use EM radiation, they provide alternative estimates that do not depend on cosmological distance 
scale~\cite{gw_distance_scale}. Information on dark energy density at $z>0.1$, past the edge of the local universe, 
could be obtained to better constrain cosmological parameters. At this distance, the cosmological distance-redshift 
relation depends on more than just $H_0$~\cite{gw_cosmology}. A further dependency is on the matter and dark energy 
densities $\Omega_m$ and $\Omega_\Lambda$. The values of these parameters could be constrained by measurements of the 
background rate of expansion from gravitational wave (GW) events at non-local redshifts. Next generation GW detectors 
such as Cosmic Explorer~\cite{cosmic_explorer} and the Einstein Telescope~\cite{einstein_telescope} as well as space 
based detectors such as LISA~\cite{lisa} will be at the forefront of such observations.
There are also gravitational waves from epochs in the early universe. These are known as relic gravitational waves and 
they can be attributed to GWs in the paradigm of quintessential inflation~\cite{relic_gws}. GW astrophysics has potential 
in the future to open new windows onto the conditions of the early universe.

@@

\subsection{Simulation data}
Hydrodynamical simulations of particles with masses on the order of a typical galaxy are the best way to emulate 
cosmological phenomena for testing without observational complications. Cosmological simulations model the physics in 
the evolution of baryonic and dark-matter in a self-consistent way that does not depend on observations.

@@

\subsection{Extraction of training data} % TRAINING DATA MANUFACTURE AND PREPROCESSING
Position vector components for the centres of cube samples in the simulation volume were drawn from a uniform distribution 
along each axis of each redshift-snapshot. If the addition or subtraction of half the side length of the cube to any of 
the components resulted in a point outside of the simulation volume, the point was discarded. The galaxy cube samples were 
obtained by sorting the position vectors of galaxies in the simulation that were within the cube volume. Each point inside 
the cube was recorded in an array. From each simulation volume of $512 \: h^{-1} \,\textup{Mpc}$ ensembles of 64 array 
samples of size $80 \: h^{-1} \,\textup{Mpc}$ were selected. The galaxy positions in each array for every cube were scaled
to the range $[-1,+1]$. 3D-histograms with 32 bins per axis were made on each cube across the same range for each axis. 
The densities in each bin cell were also scaled to this range to fit the output ranges of the GAN models. 
% After testing different sizes, a cube size of $160 \: h^{-1} \,\textup{Mpc}$ was found to balance the selection of unique 
% samples in the simulation volume against featuring the same features in the simulation structure.

@@

\textit{Architecture} --- 
Table~\ref{table:3dgan_archi} of Appendix~\ref{appendix:gan_setups} shows the architectures of the 3D-GAN discriminator 
and generator models. The hyper-parameters used in the 3D-GAN are also given in Table~\ref{table:3dgan_params} of 
Appendix~\ref{appendix:gan_setups}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. \\

{\setlength{\parindent}{0cm}
\textit{Training} --- 
The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Tensorflow~\cite{tensorflow} and 
Keras~\cite{keras}. The generator transforms a 256-dimensional vector drawn from a Gaussian prior $\mathbf{z}\sim 
\mathcal{N}(0,1)$ into a $32 \times 32 \times 32$ single-channel image. The latent-space vector is followed by a fully 
connected layerthat is reshaped to a stack of feature maps for the convolutional layers. 
% Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} 
% for more detail on batch-normalisation and activation functions.

The 3D-ACGAN implementation was limited by the GPU memory capacity - the parameters and architectures together with the 
sample batch sizes meant that the GPU was easily saturated. Deeper models with more convolutional units were implemented 
but these tended to stall the GPU. A training loop for $10^5$ epochs for the $160 \: h^{-1}\: \textup{Mpc}$ cubes 
formatted to $32 \times 32 \times 32$ pixel histograms took approximately 72 hours on an NVIDIA Tesla V100 GPU with 
32GB capacity. 
}

Generative adversarial networks are harder to train with three-dimensional input data because of the increase in the 
number of configurations of interest in the data. There are a number of tricks and techniques for moving toward more 
convergent and stable training. A compilation of these resulted from testing GANs in natural image tasks. The techniques 
that were used to train the 3D-GAN in this work were one-sided label smoothing, historical averaging and a decaying noise 
input which were borrowed from~\cite{gantricks_sali, gan_noise_decay}. The exact implementations stem only from the 
original theoretical basis. One-sided label smoothing adds scalar values sampled from a uniform distribution $\mathcal{U}
(-0.1,0)$ to the real labels only. Historical averaging adds a penalty $L_H$ to the objective functions of the discriminator
and generator. The addition is

@@

where the free parameters $A=10^{-2}$ and $\gamma=10^{-3}$ were found to work well. These techniques were not necessary in 
the 2D-GAN implementation. This is likely due to the relative information content between the 2D and 3D samples. Diluting 
the real and fake batches of samples to $1\%$ of the batch-size with the opposite validity of samples helped the training 
process stabilise.

@@

\subsubsection{Auxiliary Classification}
The ACGAN uses a secondary classification model to designate the discriminators belief in the type of image it has been 
shown. This model is nothing more than a classification neural network that is connected to the feature-extracting 
convolutions of the discriminator model. The model is optimised together with the discriminator using a sparse categorical
cross entropy function that calculates the cross entropy between two one-hot-encoded classification vectors that pertain 
to the class of an image. In practice a softmax activation function is used, which defines a discrete probability 
distribution over the discrete class labels.

@@

@@

\section{Statistics and diagnostics}\label{sec:stats_and_diags}
% WHAT ARE THESE MEASURES FOR? WHY ARE THEY IMPORTANT TO THE EXPERIMENT ? WHAT DO THE STATS CONSITS OF IN MAKING? WHAT DO THEY SHOW?
\subsection{Comparison of distributions with Kullback-Leibler Divergences}
Generative Adversarial Networks reduce the difference of the model probability distribution between the generated samples 
$p_{model}$ and the probability distribution of the training data samples $p_{data}$. This reduction quantifies how well 
$p_{model}$ estimates $p_{data}$. An estimate on the suitability of this replacement was made using a comparison with 
different ensembles of samples. The ensembles were collected from the training, generated and uniform distributions of 
histograms denoted $t$, $g$ and $u$. The uniform distribution $u$ was a set of histograms with pixel densities drawn from 
a uniform distribution $\mathcal{U}(-1,1)$. The comparison of sample ensembles tested how well the generator made the 
discriminator associate the real and generated samples to the same distribution. The Kullback-Leibler (KL) divergence and 
its relevance to the comparisons is defined in Appendix~\ref{appendix:kl_div}.

@@

@@

\subsection{Generated images}

@@

\subsection{KL trials}\label{sec:kl_trials}

@@

\subsection{Improvements to the present method}
The strongest dependence of the quality and realism of the generator product are the GAN models and hyper-parameters. It 
should be noted that each of these should really be tested independently but it is quite possible that a lifetime is not 
long enough to do so. The author's opinion is that the best route is to improve the current architecture by increasing the 
number of pixels in the cubes, allowing the depth of the models to increase with more layers, whilst also better translating
the variance in the structure inside the histograms. It has been found that in amplifying the initial parameterised units 
in the first layer of the generator model requires that there are at least $\sim \! 10^4$ units in the 3D case. The use of
fewer units does not allow the generator to draw enough permutations from the potentials  `proto-cubes' allowed by the 
initial layer. The kernels in the convolutional layers were found to be most effective at size $5 \times 5 \times 5$ in 
the 3D case and the equivalent for the 2D case. Though it is expected that some combination of higher kernel sizes would 
better characterise the histogram structures.

A larger original simulation with which to sample training data would help the networks process differences between 
individual samples. The larger volume would give more homogeneous samples with less variance in the summary statistics 
per sample. The means, total densities and variances of the pixel distributions in each cube were calculated in each 
iteration of training. Monitoring these statistics showed that the their variance was up to 10\% of their total values 
each. Increasing the homogeneity of the samples would decrease this value. The sizes of the samples relative to the 
original simulation volume could also be increased from the tested value to increase the homogeneity of the samples. 
Increasing the sample cube size would also allow the $D_{KL}$ testing to consider more varied structure in the kernel 
sizes.

The $D_{KL}$ trials could be repeated to compare the 3D two-dimensional cube projections with the 2D generated images 
and training data. This would connect the results of the separate 2D and 3D implementations. For both of the 
implementations the $p_{KS}$ testing could have been repeated in the same way as the $D_{KL}$ tests to give more 
understanding of the differences between the distributions. In this work the $p_{KS}$-values would be better understood
if they were compared to other distribution comparison $p_{KS}$-values.

@@

@@

>>>>

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents a three-dimensional generative framework that uses an Auxiliary Classifier Generative
		    Adversarial Network to create realistic samples of cosmological structure at various values of redshift~$z$. Provisional samples of the virtual cosmic web are obtained from N-body simulation data that is restricted to 
		    volumes and projections of size $200 \, h^{-1} \, \text{Mpc}$. The generated samples are shown to demonstrate
		    a high degree of coherence with their simulated counterparts and the distinction is difficult to make visually
		    by human experts. The nature of the separation between generated and N-body simulation samples is quantified 
		    with repeated tests of the Kullback-Leibler divergence on ensembles of samples alongside samples with 
		    uniformly distributed pixel densities. It is found that the implementations secure a coherence to within the
		    same order of magnitude as a self comparison of the simulation samples. The main advantage of the generative
		    approach to simulating large volumes of cosmological structure is that much less time is required than 
		    creating N-body simulations on high performance computing resources. This is demonstrated by the creation of
		    large box volumes made with $\sim10^3$ generated samples for a redshift $z$ arranged on a cubic mesh. At this 
		    stage, the generative artificial intelligence can then be said to be capable of generating an artificial 
		    universe. Future explorations of the method that are relevant to the next paradigm in cosmological simulations 
		    are also outlined. 
		  %  The most pertinent application being for a generator that can produce samples that maintain high statistical coherence to simulation data at various redshifts.

@@

\setcounter{tocdepth}{2}
\begingroup
\let\clearpage\relax
\tableofcontents
\endgroup
\newpage

\section{Introduction}
Beyond the distance scales of galactic diameters, toward the sizes of galaxy clusters and beyond, matter in the Universe 
appears as a webbed structure of filaments across volumes of emptiness known as voids. This network of luminous formation 
is known as the cosmic web~\cite{bond_cw1, Coles_cw2, Forero_cw3, Dietrich_cw4, cw_phenomenon}. The cosmic web as a 
distribution of matter contains artefacts linked to the state of the primordial universe, the nature of early universe 
perturbations and more fundamental physics to describe this era in cosmic time. The emergence of the structure is from 
an early period of exponential expansion, known as inflation~\cite{inflation_cosmo2}. This structure has been observed by 
missions such as~\cite{sdss1, 2df} and by weak gravitational lensing~\cite{glens_book} through distortions, or 
shears~\cite{shear_review}, to images of distant galaxies caused by the large-scale structure of the universe.

@@

\subsection{Generative adversarial learning}
% introduce adversarial relation / game between D and G
The discriminator and the generator are neural networks that each network try to reduce their own objective function. 
This adversarial relationship is a game between the models rather than two optimisation problems. The optimum in this 
case is a Nash equilibrium~\cite{NIPS16} rather than a local minimum given by optimisation in the case of a singular 
neural network. In this equilibrium the returns of the objective functions only change by indirect influence of the 
other network and at this stage the gradients for backpropagation to either model have vanished.

@@

\begin{equation}
    \bm{\theta}^{(G)*} = \argmin_{\mathbf{\theta}^{(G)}}\max_{\mathbf{\theta}^{(D)}}
                          V({\mathbf{\theta}^{(D)}},{\mathbf{\theta}^{(G)}})
\end{equation}

@@

% The cross-entropy for the discriminator classification of a sample against the true label of the sample is numerically 
% effective for a GAN in practice because this objective does not saturate for an incorrect identification. Because the 
% networks try to minimize their opposing objectives, if the discriminators performance increases enough, the generator 
% objective function tends to zero. This causes the generator performance to become static. At all points in the parameter 
% space the generator objective minimization is guided by the gradients from backpropagation which will reduce to zero. At 
% this point the discriminator cannot improve and less information is back-propagated through the networks.

@@

% Label embedding, one hot encoding, cross entropy

\subsection{Fourier Transform ACGAN (FT-ACGAN)}
% write this if I get it to work

@@

\subsection{Gravitational wave sources} % Rachel Gray et al. paper
Measurements of cosmological parameters from observations of gravitational wave events are independent of the 
electromagnetic (EM) spectrum and any obstacles to the transmission of EM radiation. Because these measurements 
do not use EM radiation, they provide alternative estimates that do not depend on cosmological distance 
scale~\cite{gw_distance_scale}. Information on dark energy density at $z>0.1$, past the edge of the local universe, 
could be obtained to better constrain cosmological parameters. At this distance, the cosmological distance-redshift 
relation depends on more than just $H_0$~\cite{gw_cosmology}. A further dependency is on the matter and dark energy 
densities $\Omega_m$ and $\Omega_\Lambda$. The values of these parameters could be constrained by measurements of the 
background rate of expansion from gravitational wave (GW) events at non-local redshifts. Next generation GW detectors 
such as Cosmic Explorer~\cite{cosmic_explorer} and the Einstein Telescope~\cite{einstein_telescope} as well as space 
based detectors such as LISA~\cite{lisa} will be at the forefront of such observations.
There are also gravitational waves from epochs in the early universe. These are known as relic gravitational waves and 
they can be attributed to GWs in the paradigm of quintessential inflation~\cite{relic_gws}. GW astrophysics has potential 
in the future to open new windows onto the conditions of the early universe.

@@

\subsection{Simulation data}
% Basic intro to Cosmology simulations, describe simulation being used, redshift details
Hydrodynamical simulations of particles with masses on the order of a typical galaxy are the best way to emulate 
cosmological phenomena for testing without observational complications. Cosmological simulations model the physics in 
the evolution of baryonic and dark-matter in a self-consistent way that does not depend on observations.

@@

\subsection{Extraction of training data} 
% TRAINING DATA AND PREPROCESSING
Position vector components for the centres of cube samples in the simulation volume were drawn from a uniform distribution 
along each axis of each redshift-snapshot. If the addition or subtraction of half the side length of the cube to any of 
the components resulted in a point outside of the simulation volume, the point was discarded. The galaxy cube samples were 
obtained by sorting the position vectors of galaxies in the simulation that were within the cube volume. Each point inside 
the cube was recorded in an array. From each simulation volume of $512 \: h^{-1} \,\textup{Mpc}$ ensembles of 64 array 
samples of size $80 \: h^{-1} \,\textup{Mpc}$ were selected. The galaxy positions in each array for every cube were scaled
to the range $[-1,+1]$. 3D-histograms with 32 bins per axis were made on each cube across the same range for each axis. 
The densities in each bin cell were also scaled to this range to fit the output ranges of the GAN models. 
% After testing different sizes, a cube size of $160 \: h^{-1} \,\textup{Mpc}$ was found to balance the selection of unique 
% samples in the simulation volume against featuring the same features in the simulation structure.

@@

\textit{Architecture} --- 
Table~\ref{table:3dgan_archi} of Appendix~\ref{appendix:gan_setups} shows the architectures of the ACGAN discriminator 
and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
Appendix~\ref{appendix:acgan_setups}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. \\

{\setlength{\parindent}{0cm}
\textit{Training} --- 
The networks follow the structuring of the DCGAN models~\cite{dcgan} and are written in Tensorflow~\cite{tensorflow} and 
Keras~\cite{keras}. The generator transforms a 256-dimensional vector drawn from a Gaussian prior $\mathbf{z}\sim 
\mathcal{N}(0,1)$ into a $32 \times 32 \times 32$ single-channel image. The latent-space vector is followed by a fully 
connected layer that is reshaped to a stack of feature maps for the convolutional layers. 
% Through each layer hyperbolic tangent activations are used. See Appendices~\ref{appendix:app_acts} and~\ref{app_BN} 
% for more detail on batch-normalisation and activation functions. 
}

The ACGAN implementation was limited by the GPU memory capacity - the parameters and architectures together with the 
sample batch sizes meant that the GPU was easily saturated. Deeper models with more convolutional units were implemented 
but these tended to stall the GPU. A training loop for $10^5$ epochs for the $200 \: h^{-1}\: \textup{Mpc}$ cubes 
formatted to $32 \times 32 \times 32$ pixel histograms took approximately 72 hours on an NVIDIA Tesla V100 GPU with 
32GB capacity.

Generative adversarial networks are harder to train with three-dimensional input data because of the increase in the 
number of configurations of interest in the data. There are a number of tricks and techniques for moving toward more 
convergent and stable training. A compilation of these has resulted from testing GANs in natural image tasks. The techniques
that were used to train the ACGAN in this work were one-sided label smoothing, historical averaging and a decaying noise 
input which were borrowed from~\cite{gantricks_sali, gan_noise_decay}. The exact implementations stem only from the 
original theoretical basis. One-sided label smoothing adds scalar values sampled from a uniform distribution $\mathcal{U}
(-0.1,0)$ to the real labels only. Historical averaging~\cite{gantricks_sali} adds a penalty $L_H$ to the objective functions 
of the discriminator and generator. The addition is

@@

where the free parameters $A=10^{-2}$ and $\gamma=10^{-3}$ were found to work well. Diluting the real and fake batches of 
samples to $1\%$ of the batch-size with the opposite validity of samples helped the training process stabilise.

@@

\subsubsection{Auxiliary Classification}
The ACGAN uses a secondary classification model to designate the discriminators belief in the type of image it has been 
shown. This model is nothing more than a classification neural network that is connected to the feature-extracting 
convolutions of the discriminator model. The model is optimised together with the discriminator using a sparse categorical
cross entropy function that calculates the cross entropy between two one-hot-encoded classification vectors that pertain 
to the class of an image. In practice a softmax activation function is used, which defines a discrete probability 
distribution over the class labels for each redshift value. 
% This auxiliary classifier model is optimised and trained externally to the discriminator and generator models.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generating a universe}

% {\setlength{\parindent}{0cm}
% \textit{Redshift interpolation} --- 
% To observe the A fixed point in the latent space is chosen and 
% }
\subsubsection{Redshift interpolation}
% talk about what this is in our method, what comes out of doing it 
The latent space of a generator in a typical GAN framework stores a compressed representation of the statistical structure
of the distribution of the real data. By choosing different points in this space and supplying them to the generator, 
samples can be derived with different characteristics that have been extracted from the training process of the discriminator
and generator models.

An array of latent-space vectors and class labels of a given redshift are passed to the trained ACGAN generator to create a
collection of cube samples. These are placed together face-to-face to create larger samples of cosmic structure.

\subsubsection{Latent space navigation}
% show the maths behind it, the general method of doing it. See https://arxiv.org/pdf/1609.04468.pdf, % (Larsen et al., 2016).
In all non-trivial cases, the number of parameters available to the generative model of a GAN through its architecture is less than the total parameters that constitute the training set. This means the discriminator and generator must find 
condensed representations of the training data. The generator model is sampled from a set of latent variables in the latent
space. Vector space arithmetic with points in the latent space allows interpolation between tenable examples of learned 
representations of the training data the GAN is exposed to in training. 
% With the inherent structuring of the latent space by GANs, these operations can allow complex

Interpolation is the transit between two distinct points in the latent space. In this work, spherical interpolation is used
as opposed to linear interpolation. The latter being a over-simplistic approach as the latent space of the model trained in this work had 256 dimensions with a Gaussian prior $\mathcal{N}(0,1)$. Spherical interpolation uses a great circle arc on the N-dimensional hypersphere in the latent space, with a path given by % (Shoemake 85)

\begin{equation}
    S(q_1, q_2; \mathbf{\mu}) = \frac{\sin (1 - \mu)\theta}{\sin \theta}q_1 + \frac{\sin \mu\theta}{\sin \theta}q_2
\end{equation}

where a discrete set of points between points $q_1$ and $q_2$ are passed to the generator model to create samples with.

\subsubsection{Building larger volumes from generated samples}
% say how the larger volumes are made from smaller generated cubes, talk about 
A selection of cubes were generated with a batch of $N$ latent vectors $\mathbf{Z}=\{\mathbf{z}_i, ..., \mathbf{z}_N\}$ drawn from the same prior $\mathcal{N}(0,1)$ as the generative model in training. These are placed at points on a cubic mesh to create a larger box of generated cubes. This was repeated for side-lengths 7, 9 and 11 (in multiples of the initial generated cube side-lengths).

@@

\section{Statistics and diagnostics}\label{sec:stats_and_diags}
% WHAT ARE THESE MEASURES FOR? WHY ARE THEY IMPORTANT TO THE EXPERIMENT ? 
% WHAT DO THE STATS CONSIST OF IN MAKING? WHAT DO THEY SHOW?

\subsection{Comparison of distributions with Kullback-Leibler Divergences}
Generative Adversarial Networks reduce the difference of the model probability distribution between the generated samples 
$p_{model}$ and the probability distribution of the training data samples $p_{data}$. This reduction quantifies how well 
$p_{model}$ estimates $p_{data}$. An estimate on the suitability of this replacement was made using a comparison with 
different ensembles of samples. The ensembles were collected from the training, generated and uniform distributions of 
histograms denoted $t$, $g$ and $u$. The uniform distribution $u$ was a set of histograms with pixel densities drawn from 
a uniform distribution $\mathcal{U}(-1,1)$. The comparison of sample ensembles tested how well the generator made the 
discriminator associate the real and generated samples to the same distribution. The Kullback-Leibler (KL) divergence and 
its relevance to the comparisons is defined in Appendix~\ref{appendix:kl_div}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Power spectrum correlation}
% I have a code to generate power spectra for a set of cubes. compare training / generated / uniform spectra?

@@

\subsection{Generated images}
% show generated images at various redshifts, compare with training data

\begin{figure*}[hbt!]
\includegraphics[width=15cm]{figures/cubes/classes3d.jpg}
\centering
\caption{A collection of training data cubes extracted from the Millennium-SAGE catalogs. Each row corresponds to the redshift value given above each cube. The scaling $s(x)$ is applied to each cube $x$.}
\label{fig:gan_game}
\end{figure*}

\subsubsection{Redshift interpolation}
% show grid of cubes ranging from z=0 to z=(highest z in training data classes)

\subsubsection{Universe generation}
% show three scales (side lengths in terms of generated cubes - e.g. 7, 9, 11) of generated-cube-universes.

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/mill_box_3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.19$}
\label{fig:kernel_fig}
\end{figure}

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/gen_box_9_z3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.0$.}
\label{fig:kernel_fig}
\end{figure}

@@

\subsection{KL trials}\label{sec:kl_trials}
% do the same as before (in thesis) but average over classes?

@@

\subsection{Improvements to the present method}
The strongest dependence of the quality and realism of the generator product are the GAN models and hyper-parameters. It 
should be noted that each of these should really be tested independently but it is quite possible that a lifetime is not 
long enough to do so. The author's opinion is that the best route is to improve the current architecture by increasing the 
number of pixels in the cubes, allowing the depth of the models to increase with more layers, whilst also better translating
the variance in the structure inside the histograms. It has been found that in amplifying the initial parameterised units 
in the first layer of the generator model requires that there are at least $\sim \! 10^4$ units in the 3D case. The use of
fewer units does not allow the generator to draw enough permutations from the potentials  `proto-cubes' allowed by the 
initial layer. The kernels in the convolutional layers were found to be most effective at size $5 \times 5 \times 5$. 
Though it is expected that some combination of higher kernel sizes would better characterise the histogram structures.

A larger original simulation with which to sample training data would help the networks process differences between 
individual samples. The larger volume would give more homogeneous samples with less variance in the summary statistics 
per sample. 
% The means, total densities and variances of the pixel distributions in each cube were calculated in each 
% iteration of training. Monitoring these statistics showed that the their variance was up to 10\% of their total values 
% each. Increasing the homogeneity of the samples would decrease this value. 
The sizes of the samples relative to the original simulation volume could also be increased from the tested value to
increase the homogeneity of the samples. Increasing the sample cube size would also allow the $D_{KL}$ testing to consider
more varied structure in the kernel sizes.

% The $D_{KL}$ trials could be repeated to compare the 3D two-dimensional cube projections with the 2D generated images 
% and training data. This would connect the results of the separate 2D and 3D implementations. For both of the 
% implementations the $p_{KS}$ testing could have been repeated in the same way as the $D_{KL}$ tests to give more 
% understanding of the differences between the distributions. In this work the $p_{KS}$-values would be better understood
% if they were compared to other distribution comparison $p_{KS}$-values.

@@

% Find a way to co-interpolate a massive volume of cubes between different redshifts

@@

\subsection{Structuring the latent space}
% as in INFOGAN. I might be able to do this for this ACGAN. Would be really good.
